---
title: "Problem_1"
author: "Pol Parra"
date: "2024-03-22"
output: html_document
---



```{r}
install.packages("openxlsx")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
```


```{r}
train_data <- read.csv("train.csv")
head(train_data)
```

1. Select only the variables of interest.
2. Transform the response variable so that you have Saleprice in  tens of thousands of dollars.
     
```{r}
variables_of_interest <- train_data[, c("SalePrice", "CentralAir", "KitchenQual")]
variables_of_interest$SalePrice <- variables_of_interest$SalePrice / 10000
head(variables_of_interest)
```


3. Consider only the levels "Gd", "Fa" and "TA" in the variable "kitchenqual", transforming the level "Ex" to "Gd".

KitchenQual: Kitchen quality

       Ex	Excellent
       Gd	Good
       TA	Typical/Average
       Fa	Fair
       Po	Poor

```{r}
variables_of_interest$KitchenQual <- ifelse(variables_of_interest$KitchenQual == "Ex", "Gd", variables_of_interest$KitchenQual)
variables_of_interest <- subset(variables_of_interest, KitchenQual %in% c("Gd", "Fa", "TA"))
head(variables_of_interest)
```

print(variables_of_interest)



4. In order to have a balanced analysis, you randomly select 6 observations from the different groups.

--> Observations: Y/N central Air * 3 Groups (Ex, Gd, TA) * 6 observations each = 36 items




```{r}
library(dplyr)
balanced_sample <- variables_of_interest %>%
  group_by(CentralAir, KitchenQual) %>%
  sample_n(6, replace = TRUE) %>% 
  ungroup() 

View(balanced_sample)
```


Exporting the data balanced to Excel in order to start our ANOVA - Regression Study.

```{r}
library(openxlsx)
write.xlsx(balanced_sample, "/Users/polparracatalan/Downloads/house-prices-advanced-regression-techniques/balanced_sample.xlsx")
```










